---
title: "Untitled"
author: "Harry Robertson"
date: '2024-03-04'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
library(tidyverse)
library(TOP)
library(directPA)
```

# Figure 2

## Identification of a pan-organ rejection signal across solid organ transplantation. 

### Figure 2A

Venn diagram showing the overlap and uniqueness of differentially expressed genes between biopsy samples from solid organ transplant recipients experiencing acute rejection and those from stably functioning grafts. The number of overlapping genes and (number of genes expected by chance)

```{r}
setwd("~/Downloads/Rejection/Processed_Datasets")
files <- list.files(pattern = "Biopsy")[-1]
Biopsy_Counts <- lapply(files, function(x){
  y <- read.csv(x) 
})
colnames(Biopsy_Counts[[10]])[2] <- "Gene.Symbol"
Biopsy_Counts <- lapply(Biopsy_Counts, function(x){
  x %>%
    dplyr::select(-X) %>%
    tibble::remove_rownames() %>%
    filter(!duplicated(Gene.Symbol) & !is.na(Gene.Symbol)) %>%
    tibble::column_to_rownames("Gene.Symbol") %>%
    dplyr::select(t)
})

keep <- Biopsy_Counts %>% sapply(nrow) > 8000
Biopsy_Counts <- Biopsy_Counts[keep]

int <- Reduce(intersect, lapply(Biopsy_Counts, rownames))
test_list <-  lapply(Biopsy_Counts, function(x)x[int,, drop = FALSE])

test_df <- do.call(cbind, test_list)
Z.Scores.All <- apply(test_df, 2, function(x) {
        stats::qnorm(rank(x)/(nrow(test_df) + 1))
})
data(Pathways, package = "directPA")
gene.pvalues <- apply(Z.Scores.All, 1, function(x) {
  directPA::geneStats(x)
})
gene.zscores <- stats::qnorm(gene.pvalues, lower.tail = FALSE)
pvalue2sided <- 2 * stats::pnorm(-abs(gene.zscores))
sig.genes <- names(sort(pvalue2sided[pvalue2sided < 1*10^-35]))

# Plotting Results
GSE <- paste(TOP:::str_split_n(files[keep], "_", 2), sub(".csv", "", TOP:::str_split_n(files[keep], "_", 4)), sep = "_")
Tissue <- sub(".csv", "", TOP:::str_split_n(files[keep], "_", 5))
Tissue[Tissue == ""] <- "Kidney"
anno_df <- data.frame(GSE, Tissue, row.names = paste("Dataset", 1:length(Tissue)))

Z.Score.Significant <- Z.Scores.All[which(rownames(Z.Scores.All) %in% sig.genes),]
colnames(Z.Score.Significant) <- rownames(anno_df)
```

```{r}
Kidney_idx <- Tissue == "Kidney"
Liver_idx <- Tissue == "Liver"
Lung_idx <- Tissue == "Lung"
Heart_idx <- Tissue == "Heart"

test_df.Kidney <- test_df[,Kidney_idx]
test_df.Liver <- test_df[,Liver_idx]
test_df.Lung <- test_df[,Lung_idx]
test_df.Heart <- test_df[,Heart_idx]

Gene_sets <- lapply(list(test_df.Kidney, test_df.Liver, test_df.Lung, test_df.Heart), function(x){
  Z.Scores.All <- apply(x, 2, function(y) {
        stats::qnorm(rank(y)/(nrow(x) + 1))
  })
  data(Pathways, package = "directPA")
  gene.pvalues <- apply(Z.Scores.All, 1, function(y) {
    directPA::geneStats(y)
  })
  gene.zscores <- stats::qnorm(gene.pvalues, lower.tail = FALSE)
  return(gene.zscores)
})

names(Gene_sets) <- c("Kidney", "Liver", "Lung", "Heart")

df <- do.call(cbind, lapply(Gene_sets, tibble::enframe))
df <- df[,grep("value", colnames(df))]
rownames(df) <- names(Gene_sets[[1]])
df <- df[!is.infinite(rowMeans(df)),]

library(GGally)
library(ggfortify)

df_venn <- abs(df) > 2 
df_venn %>%
  data.frame()

chi_square <- chisq.test(apply(df_venn, 2, table))
chi_square$residuals

tab <- table(df_venn[,1], df_venn[,2], df_venn[,3], df_venn[, 4])
tab <- as.data.frame(tab)

library(ggvenn)
df_venn %>%
  data.frame() %>%
  ggplot() +
  geom_venn(aes(A = Kidney.value, B = Liver.value, C = Lung.value, D = Heart.value),show_percentage = F,
            set_names = c("Kidney", "Liver", "Lung", "Heart")) + theme_void()
```


```{r}
library(UpSetR)
upset_df <- df_venn %>%
  data.frame() 
upset_df[] <- lapply(upset_df, as.numeric)
UpSetR::upset(upset_df, sets = colnames(upset_df))

tab <- table(upset_df)

marginal1 <- rowSums(tab) / sum(tab)
marginal2 <- colSums(tab) / sum(tab)
marginal3 <- apply(tab, 3, sum) / sum(tab)
marginal4 <- apply(tab, 4, sum) / sum(tab)

prob <- array(0, dim = dim(tab))
for (i in 1:dim(tab)[1]) {
  for (j in 1:dim(tab)[2]) {
    for (k in 1:dim(tab)[3]) {
      for (l in 1:dim(tab)[4]) {
        prob[i,j,k,l] <- marginal1[i] * marginal2[j] * marginal3[k] * marginal4[l]
      }
    }
  }
}
exp <- prob*nrow(upset_df)
residuals <- (tab - exp) / sqrt(exp)
residuals_df <- residuals %>%
  data.frame()
```

### Figure 2B

Heatmap of the top 50 rejection-specific genes, with each column representing a dataset and each row a gene.

```{r}
colnames(Z.Score.Significant) <- GSE
library(pheatmap)
pheatmap::pheatmap(Z.Score.Significant,
         color = colorRampPalette(c("blue", "navy", "white", "firebrick3", "red"))(1500),
         cellwidth = 8,
         cellheight = 10,
         angle_col = 315,
         show_colnames = FALSE)

```
