---
title: "Figure 2"
author: "Harry Robertson"
date: '2023-03-04'
output: 
  html_document:
    code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
library(tidyverse)
library(TOP)
library(directPA)
```

# Figure 2

## Identification of a pan-organ rejection signal across solid organ transplantation.

### Figure 2A

Venn diagram showing the overlap and uniqueness of differentially expressed genes between biopsy samples from solid organ transplant recipients experiencing acute rejection and those from stably functioning grafts. The number of overlapping genes and (number of genes expected by chance)

```{r}
setwd("~/Downloads/Rejection/Processed_Datasets")
files <- list.files(pattern = "Biopsy")[-1]
Biopsy_Counts <- lapply(files, function(x){
  y <- read.csv(x) 
})
colnames(Biopsy_Counts[[10]])[2] <- "Gene.Symbol"
Biopsy_Counts <- lapply(Biopsy_Counts, function(x){
  x %>%
    dplyr::select(-X) %>%
    tibble::remove_rownames() %>%
    filter(!duplicated(Gene.Symbol) & !is.na(Gene.Symbol)) %>%
    tibble::column_to_rownames("Gene.Symbol") %>%
    dplyr::select(t)
})

keep <- Biopsy_Counts %>% sapply(nrow) > 8000
Biopsy_Counts <- Biopsy_Counts[keep]

int <- Reduce(intersect, lapply(Biopsy_Counts, rownames))
test_list <-  lapply(Biopsy_Counts, function(x)x[int,, drop = FALSE])

test_df <- do.call(cbind, test_list)
Z.Scores.All <- apply(test_df, 2, function(x) {
        stats::qnorm(rank(x)/(nrow(test_df) + 1))
})
data(Pathways, package = "directPA")
gene.pvalues <- apply(Z.Scores.All, 1, function(x) {
  directPA::geneStats(x)
})
gene.zscores <- stats::qnorm(gene.pvalues, lower.tail = FALSE)
pvalue2sided <- 2 * stats::pnorm(-abs(gene.zscores))
sig.genes <- names(sort(pvalue2sided[pvalue2sided < 1*10^-35]))

# Plotting Results
GSE <- paste(TOP:::str_split_n(files[keep], "_", 2), sub(".csv", "", TOP:::str_split_n(files[keep], "_", 4)), sep = "_")
Tissue <- sub(".csv", "", TOP:::str_split_n(files[keep], "_", 5))
Tissue[Tissue == ""] <- "Kidney"
anno_df <- data.frame(GSE, Tissue, row.names = paste("Dataset", 1:length(Tissue)))

Z.Score.Significant <- Z.Scores.All[which(rownames(Z.Scores.All) %in% sig.genes),]
colnames(Z.Score.Significant) <- rownames(anno_df)
```

```{r}
Kidney_idx <- Tissue == "Kidney"
Liver_idx <- Tissue == "Liver"
Lung_idx <- Tissue == "Lung"
Heart_idx <- Tissue == "Heart"

test_df.Kidney <- test_df[,Kidney_idx]
test_df.Liver <- test_df[,Liver_idx]
test_df.Lung <- test_df[,Lung_idx]
test_df.Heart <- test_df[,Heart_idx]

Gene_sets <- lapply(list(test_df.Kidney, test_df.Liver, test_df.Lung, test_df.Heart), function(x){
  Z.Scores.All <- apply(x, 2, function(y) {
        stats::qnorm(rank(y)/(nrow(x) + 1))
  })
  data(Pathways, package = "directPA")
  gene.pvalues <- apply(Z.Scores.All, 1, function(y) {
    directPA::geneStats(y)
  })
  gene.zscores <- stats::qnorm(gene.pvalues, lower.tail = FALSE)
  return(gene.zscores)
})

names(Gene_sets) <- c("Kidney", "Liver", "Lung", "Heart")

df <- do.call(cbind, lapply(Gene_sets, tibble::enframe))
df <- df[,grep("value", colnames(df))]
rownames(df) <- names(Gene_sets[[1]])
df <- df[!is.infinite(rowMeans(df)),]

library(GGally)
library(ggfortify)

df_venn <- abs(df) > 2 

chi_square <- chisq.test(apply(df_venn, 2, table))

tab <- table(df_venn[,1], df_venn[,2], df_venn[,3], df_venn[, 4])
tab <- as.data.frame(tab)

library(ggvenn)
df_venn %>%
  data.frame() %>%
  ggplot() +
  geom_venn(aes(A = Kidney.value, B = Liver.value, C = Lung.value, D = Heart.value),show_percentage = F,
            set_names = c("Kidney", "Liver", "Lung", "Heart")) + theme_void()
```

```{r}
library(UpSetR)
upset_df <- df_venn %>%
  data.frame() 
upset_df[] <- lapply(upset_df, as.numeric)
UpSetR::upset(upset_df, sets = colnames(upset_df))

tab <- table(upset_df)

marginal1 <- rowSums(tab) / sum(tab)
marginal2 <- colSums(tab) / sum(tab)
marginal3 <- apply(tab, 3, sum) / sum(tab)
marginal4 <- apply(tab, 4, sum) / sum(tab)

prob <- array(0, dim = dim(tab))
for (i in 1:dim(tab)[1]) {
  for (j in 1:dim(tab)[2]) {
    for (k in 1:dim(tab)[3]) {
      for (l in 1:dim(tab)[4]) {
        prob[i,j,k,l] <- marginal1[i] * marginal2[j] * marginal3[k] * marginal4[l]
      }
    }
  }
}
exp <- prob*nrow(upset_df)
residuals <- (tab - exp) / sqrt(exp)
residuals_df <- residuals %>%
  data.frame()
```

### Figure 2B

Heatmap of the top 50 rejection-specific genes, with each column representing a dataset and each row a gene.

```{r}
colnames(Z.Score.Significant) <- GSE
library(pheatmap)
pheatmap::pheatmap(Z.Score.Significant,
         color = colorRampPalette(c("blue", "navy", "white", "firebrick3", "red"))(1500),
         cellwidth = 8,
         cellheight = 10,
         angle_col = 315,
         show_colnames = FALSE)
```

### Figure 2C-F

```{r}
suppressPackageStartupMessages({
  library(SingleCellExperiment)
  library(Cepo)
  library(ggplot2)
  library(dplyr)
  library(ggpubr)
})
```

```{r}
goi_mat = data.frame(gene = names(pvalue2sided), z = gene.zscores, p = pvalue2sided)
goi_mat2 = goi_mat[is.infinite(goi_mat$z),]
goi_mat = goi_mat[!is.infinite(goi_mat$z),]
goi_mat = goi_mat[order(goi_mat$z, decreasing = TRUE),]
goi_mat$gene[1:50]

goi_rank = rank(goi_mat$p)
names(goi_rank) = goi_mat$gene

idx_gene = goi_mat$p < 1*10^-35 
goi = goi_mat$gene[idx_gene]
length(goi)
```

```{r}
f = list.files("~/Desktop/For_Hani")

seuList = lapply(f, function(x) readRDS(paste0("~/Desktop/For_Hani/", x)))
seuList = unlist(seuList, recursive = FALSE)

geneIdx = Reduce(intersect, lapply(seuList, function(x) toupper(rownames(x@assays$RNA@counts))))

counter <- 0
sceList = lapply(seuList, function(x) {
  
  counter <<- counter + 1 
  print(counter)
  sce = SingleCellExperiment::SingleCellExperiment(assay = list(counts = x@assays$RNA@counts),
                                                   colData = x@meta.data)
  rownames(sce) = toupper(rownames(x@assays$RNA@counts))
  sce = sce[geneIdx,]
  cellIdx = colSums(counts(sce) == 0)/nrow(sce)
  sce = sce[,cellIdx < 1]
  sce = scater::logNormCounts(sce)
  return(sce)
  
})

sapply(sceList, function(x) {
  unique(x$Condition)
})

# minor 
sapply(sceList, function(x) {
  idx = grep("predicted", colnames(colData(x)), value = TRUE)
  idx = grep("l2|level_3", idx, value = TRUE)
  idx = idx[!grepl("score", idx)]
  table(colData(x)[, idx])
})

# major
sapply(sceList, function(x) {
  idx = grep("predicted", colnames(colData(x)), value = TRUE)
  idx = grep("l1|level_2", idx, value = TRUE)
  idx = idx[!grepl("score", idx)]
  table(colData(x)[, idx])
})

counter <- 0
cepoList = lapply(sceList, function(x) {
  
  counter <<- counter + 1
  print(counter)
  idx = grep("predicted", colnames(colData(x)), value = TRUE)
  idx = grep("l2|level_3", idx, value = TRUE)
  idx = idx[!grepl("score", idx)]
  
  idx2 = grep("predicted", colnames(colData(x)), value = TRUE)
  idx2 = grep("l1|level_2", idx2, value = TRUE)
  idx2 = idx2[!grepl("score", idx2)]
  
  colData(x)$celltype_minor = colData(x)[, idx]
  colData(x)$celltype_major = colData(x)[, idx2]
  colData(x)$celltype = paste0(colData(x)$celltype_major, "___", colData(x)$celltype_minor, "___", colData(x)$Condition)
  
  cty = names(table(x$celltype))[which(table(x$celltype) > 10)]
  #print(length(cty))
  x_subset = x[, x$celltype %in% cty]
  
  res = Cepo::Cepo(logcounts(x_subset),
                   x_subset$celltype,
                   exprsPct = 0.05, minCells = 10)
  return(res)
  
})
```


```{r}
#top 200-45 or 100-39
topGenes = unique(unlist(lapply(cepoList, function(x) unique(unlist(Cepo::topGenes(x, n = 100))))))
length(topGenes)
table(topGenes %in% goi)

cepoMat = lapply(cepoList, function(x) {
  idx = intersect(rownames(x$stats), topGenes)
  return(x$stats[idx,])
})

lapply(cepoMat, function(x) table(rownames(x) %in% goi))

lapply(c(1*10^-10, 0.01, 0.05), function(p_cut) {
  
  tmp = goi_mat[goi_mat$z > 0, ]
  idx_gene = tmp$p < p_cut
  goi1 = tmp$gene[idx_gene]
  
  tmp = goi_mat[goi_mat$z < 0, ]
  idx_gene = tmp$p < p_cut
  goi2 = tmp$gene[idx_gene]
  
  print(c(length(goi1), length(goi2)))
  
})

pList = lapply(c(1*10^-10), function(p_cut) {
  
  idx_gene = goi_mat$p < p_cut & goi_mat$z > 0
  goi1 = goi_mat$gene[idx_gene]
  idx_gene = goi_mat$p < p_cut & goi_mat$z < 0
  goi2 = goi_mat$gene[idx_gene]
  
  matEnrich = do.call(rbind, lapply(1:length(cepoMat), function(x) {
    
    mat = do.call(rbind, lapply(1:ncol(cepoMat[[x]]), function(y) {
      
      geneStats = rank(cepoMat[[x]][,y])
      #geneIdx = intersect(names(goi_rank), names(test))
      
      p.in.up <- intersect(names(geneStats), goi1)
      p.out.up <- setdiff(names(geneStats), goi1)
      
      p.in.down <- intersect(names(geneStats), goi2)
      p.out.down <- setdiff(names(geneStats), goi2)
      
      wilcoxTest.mat <- matrix("NA", ncol = 4, nrow = 2)
      colnames(wilcoxTest.mat) <- c("cty", "pvalue", "# of substrates",
                                    "dir")
      
      wilcoxTest.mat[1, 1] <- colnames(cepoMat[[x]])[[y]]
      wilcoxTest.mat[1, 2] <- stats::wilcox.test(geneStats[p.in.up],
                                                 geneStats[p.out.up], alternative = "greater")$p.value
      wilcoxTest.mat[1, 3] <- length(p.in.up)
      wilcoxTest.mat[1, 4] <- "up"
      
      wilcoxTest.mat[2, 1] <- colnames(cepoMat[[x]])[[y]]
      wilcoxTest.mat[2, 2] <- stats::wilcox.test(geneStats[p.in.down],
                                                 geneStats[p.out.down], alternative = "greater")$p.value
      wilcoxTest.mat[2, 3] <- length(p.in.down)
      wilcoxTest.mat[2, 4] <- "down"
      
      
      res = as.data.frame(wilcoxTest.mat[,1:4, drop = FALSE])
      res$dataset = x
      res$condition = unique(sceList[[x]]$Condition)
      return(res)
      
    }))
    mat = as.data.frame(mat)
    mat$pvalue = as.numeric(mat$pvalue)
    return(mat)
    
  }))
  
  matEnrich$cty[-log10(matEnrich$pvalue) > 2]
  matEnrich$interaction = interaction(matEnrich$cty, matEnrich$condition)
  matEnrich$major = sapply(strsplit(matEnrich$cty, "___"), "[[", 1)
  matEnrich$minor = sapply(strsplit(matEnrich$cty, "___"), "[[", 2)
  
  #matEnrich$major[matEnrich$minor %in% c("T", "B", "NK", "Natural.Killer.T")] = "Lymphoid"
  matEnrich$logpvalue = -log10(matEnrich$pvalue)
  matEnrich$group = paste0(matEnrich$major, "_", matEnrich$dir)
  
  p = ggplot(matEnrich, aes(x = as.factor(dir), 
                            y = -log10(pvalue), fill = as.factor(condition))) + 
    ggtitle(p_cut) + 
    geom_boxplot(show.legend = TRUE, notch = FALSE) + facet_wrap(~major, ncol = 8)
  
  p = ggboxplot(matEnrich, x = "condition", y= "logpvalue", facet.by = "group", fill = "condition",
            add = "jitter", palette = "npg") + 
    stat_compare_means(aes(method = "wilcox.test", label = paste0("p = ", after_stat(p.format))))
  
  return(p)
  
})

ptosave = pList[[1]]
ptosave %>% ggsave(filename = "boxplot_enrichment.pdf", width = 15, height = 25)
```

```{r}
pList = lapply(c(1*10^-10), function(p_cut) {
  
  idx_gene = goi_mat$p < p_cut & goi_mat$z > 0
  goi1 = goi_mat$gene[idx_gene]
  idx_gene = goi_mat$p < p_cut & goi_mat$z < 0
  goi2 = goi_mat$gene[idx_gene]
  
  tmp = do.call(rbind, lapply(1:length(cepoMat), function(x) {
    
    mat = do.call(rbind, lapply(1:ncol(cepoMat[[x]]), function(y) {
      
      geneStats = rank(cepoMat[[x]][,y])
      #geneIdx = intersect(names(goi_rank), names(test))
      
      p.in.up <- intersect(names(geneStats), goi1)
      p.out.up <- setdiff(names(geneStats), goi1)
      
      p.in.down <- intersect(names(geneStats), goi2)
      p.out.down <- setdiff(names(geneStats), goi2)
      
      wilcoxTest.mat <- matrix("NA", ncol = 4, nrow = 2)
      colnames(wilcoxTest.mat) <- c("cty", "pvalue", "# of substrates",
                                    "dir")
      
      wilcoxTest.mat[1, 1] <- colnames(cepoMat[[x]])[[y]]
      wilcoxTest.mat[1, 2] <- stats::wilcox.test(geneStats[p.in.up],
                                                 geneStats[p.out.up], alternative = "greater")$p.value
      wilcoxTest.mat[1, 3] <- length(p.in.up)
      wilcoxTest.mat[1, 4] <- "up"
      
      wilcoxTest.mat[2, 1] <- colnames(cepoMat[[x]])[[y]]
      wilcoxTest.mat[2, 2] <- stats::wilcox.test(geneStats[p.in.down],
                                                 geneStats[p.out.down], alternative = "greater")$p.value
      wilcoxTest.mat[2, 3] <- length(p.in.down)
      wilcoxTest.mat[2, 4] <- "down"
      
      
      res = as.data.frame(wilcoxTest.mat[,1:4, drop = FALSE])
      res$dataset = x
      res$condition = unique(sceList[[x]]$Condition)
      return(res)
      
    }))
    mat = as.data.frame(mat)
    mat$pvalue = as.numeric(mat$pvalue)
    return(mat)
    
  }))
  
  matEnrich = tmp
  matEnrich$interaction = interaction(matEnrich$cty, matEnrich$condition)
  matEnrich$major = sapply(strsplit(matEnrich$cty, "___"), "[[", 1)
  matEnrich$minor = sapply(strsplit(matEnrich$cty, "___"), "[[", 2)
  
  #matEnrich$major[matEnrich$minor %in% c("T", "B", "NK", "Natural.Killer.T")] = "Lymphoid"
  matEnrich$logpvalue = -log10(matEnrich$pvalue)
  matEnrich = matEnrich[matEnrich$major %in% c("Immune", "Lymphoid", "Myeloid"), ] 
  matEnrich = matEnrich[!matEnrich$minor %in% c("Inner.Medullary.Collecting.Duct", "AT2", "Plasmacytoid.Dendritic", "Plasma"), ] 
  
  matEnrich$minor[matEnrich$minor %in% c("B")] = "B.cell.lineage"
  matEnrich$minor[matEnrich$minor %in% c("T")] = "T.cell.lineage"
  matEnrich$minor[matEnrich$minor %in% c("M2.Macrophage", "Macrophages")] = "Macrophage"
  matEnrich$minor[matEnrich$minor %in% c("Mast.cells")] = "Mast"
  matEnrich$minor[matEnrich$minor %in% c("Natural.Killer.T", 
                                         "Innate.lymphoid.cell.NK")] = "NK"
  matEnrich$minor[matEnrich$minor %in% c("Monocyte.cDC", "Classical.Dendritic", "Non.classical.monocyte")] = "Dendritic.cells"
  matEnrich$minor[matEnrich$minor %in% c("Monocyte.derived")] = "Monocytes"

  matEnrich$group = paste0(matEnrich$major, "_", matEnrich$minor, "_", matEnrich$dir)
    
  p = ggplot(matEnrich, aes(x = as.factor(dir), 
                            y = -log10(pvalue), fill = as.factor(condition), col = as.factor(condition))) + 
    #ggtitle(p_cut) + 
    geom_boxplot(show.legend = TRUE, notch = FALSE, alpha = 0.5) +
    #geom_jitter() + 
    ggsci::scale_fill_lancet() +
    ggsci::scale_color_lancet() +
    facet_wrap(~minor+major, ncol = 14) + 
    theme_bw()
  
  p1 = ggboxplot(matEnrich, x = "condition", y= "logpvalue", facet.by = "group", fill = "condition", ncol = 12,
            add = "jitter", palette = "npg") + 
    stat_compare_means(aes(method = "wilcox.test", label = paste0("p = ", after_stat(p.format))))
  
  return(p)
  
})

ptosave = pList[[1]]
ptosave %>% ggsave(filename = "boxplot_enrichment_minor.pdf", width = 15, height = 5)
```

```{r}
rm(seuList, sceList)
gc()

#idx_control = unlist(lapply(sceList, function(x) unique(x$Condition))) == "Control"
topGenes2 = Reduce(intersect, (lapply(cepoList, function(x) unique(unlist(Cepo::topGenes(x, n = 1000))))))
topGenes2 = unique(unlist(lapply(cepoList, function(x) unique(unlist(Cepo::topGenes(x, n = 50))))))
length(topGenes2)
cepoMat2 = lapply(cepoList, function(x) {
  idx = intersect(rownames(x$stats), topGenes2)
  return(x$stats[idx,])
})
idx = Reduce(intersect, lapply(cepoMat2, rownames))
length(idx)

cepoMat_combined = do.call(cbind, lapply(cepoMat2, function(x) x[idx,]))
corMat = cor(as.matrix(cepoMat_combined))
colnames(corMat) = paste0(colnames(corMat), "___", rep(1:length(cepoMat2), sapply(cepoMat2, ncol)))
corMat = corMat[!grepl("None|Neuronal", colnames(corMat)), !grepl("None|Neuronal", colnames(corMat))]
annot = data.frame(
  minor = sapply(strsplit(colnames(corMat), "___"), "[[", 2),
  major = sapply(strsplit(colnames(corMat), "___"), "[[", 1),
  condition = sapply(strsplit(colnames(corMat), "___"), "[[", 3)
  #sample = sapply(strsplit(colnames(corMat), "___"), "[[", 4)
)

# major
annot$major[annot$minor %in% c("T", "B", "NK", "Natural.Killer.T")] = "Immune"
annot$major[annot$major %in% c("Mast", "Lymphoid", "Myeloid")] = "Immune"
annot$major[annot$major %in% c("Proximal.Tubule", "Connecting.Tubule", "Distal.Convoluted.Tubule", 
                               "Ascending.Thin.Limb", "Descending.Thin.Limb","Thick.Ascending.Limb",
                               "Papillary.Tip.Epithelial", "Parietal.Epithelial", "Intercalated",
                               "Pericyte", "Podocyte", "Principal")] = "Kidney"
annot$major[annot$major %in% c("Alveolar.epithelium", "Airway.epithelium")] = "Lung"
annot$major[annot$major %in% c("Smooth.muscle", "Smooth.Muscle", "Vascular.Smooth.Muscle...Pericyte",
                               "Fibroblast.lineage", "Mesothelium", "Mesothelial",
                               "Lymphatic.Endothelial", "Lymphatic.EC", "Endothelial",
                               "Blood.vessels", "Fibroblast", "Adipocyte")] = "Stromal"
annot$major[annot$major %in% c("Cardiomyocyte")] = "Heart"

# minor
annot$minor[annot$minor %in% c("B")] = "B.cell.lineage"
annot$minor[annot$minor %in% c("T", "Natural.Killer.T")] = "T.cell.lineage"
annot$minor[annot$minor %in% c("Afferent...Efferent.Arteriole.Endothelial", "Arterial.Endothelial", "EC.arterial", "EC.capillary", "EC.venous", "Venous.Endothelial", "Capillary.Endothelial")] = "Endothelial-Vascular"
annot$minor[annot$minor %in% c("Glomerular.Capillary.Endothelial", "Ascending.Vasa.Recta.Endothelial", "Descending.Vasa.Recta.Endothelial.","Peritubular.Capilary.Endothelial.")] = "Endothelial-Kidney"
annot$minor[annot$minor %in% c("Lymphatic.EC.differentiating", "Lymphatic.EC.mature", "Lymphatic.Endothelial")] = "Endothelial-Lymphatic"
annot$minor[annot$minor %in% c("Macrophages", "M2.Macrophage")] = "Macrophage"
annot$minor[annot$minor %in% c("Monocyte.cDC", "Monocyte.derived", "Monocytes", "Non.classical.monocyte")] = "Monocyte"
annot$minor[annot$minor %in% c("AT1", "AT2")] = "AT"
annot$minor[annot$minor %in% c("Innate.lymphoid.cell.NK")] = "NK"
annot$minor[annot$minor %in% c("SM.activated.stress.response")] = "Smooth.Muscle"
annot$minor[annot$minor %in% c("Cortical.Collecting.Duct.Intercalated.Type.A", "Cortical.Collecting.Duct.Principal",
                               "Cortical.Collecting.Duct.Principal", "Connecting.Tubule",
                               "Cortical.Collecting.Duct.Intercalated.Type.A",
                               "Cortical.Collecting.Duct.Principal", "Inner.Medullary.Collecting.Duct",
                               "Outer.Medullary.Collecting.Duct.Intercalated.Type.A",
                               "Distal.Convoluted.Tubule")] = "Collecting duct/tubule"
annot$minor[annot$minor %in% c("Cortical.Thick.Ascending.Limb",
                               "Ascending.Thin.Limb",
                               "Descending.Thin.Limb.Type.1",
                               "Descending.Thin.Limb.Type.2",
                               "Descending.Thin.Limb.Type.3",
                               "Medullary.Thick.Ascending.Limb")] = "Descending/Ascending limb"
annot$minor[annot$minor %in% c("Proximal.Tubule.Epithelial.Segment.1",
                               "Proximal.Tubule.Epithelial.Segment.2",
                               "Proximal.Tubule.Epithelial.Segment.3")] = "Proximal.Tubule.Epithelial.Segment"
annot$minor[annot$minor %in% c("Intercalated.Type.B", "Proximal.Tubule.Epithelial.Segment", "Papillary.Tip.Epithelial", "Parietal.Epithelial")] = "Epithelial-Kidney"
annot$minor[annot$minor %in% c("Dendritic.cells", "Classical.Dendritic")] = "Dendritic cells"
annot$minor[annot$minor %in% c("Fibroblasts", "Medullary.Fibroblast", "Myofibroblasts")] = "Fibroblast"
annot$minor[annot$minor %in% c("Mast.cells")] = "Mast"
annot$minor[annot$minor %in% c("Mesangial", "Vascular.Smooth.Muscle...Pericyte", "Dendritic cells",
                               "Cortical.Vascular.Smooth.Muscle...Pericyte",
                               "Endocardial", "Adipocyte", "Pericyte")] = "Stromal"

annot$major[annot$minor == "Ventricular.Cardiomycoyte"] = "Heart"
annot$major[annot$minor == "AT"] = "Lung"
annot$major[annot$minor == "Collecting duct/tubule"] = "Kidney"
annot$major[annot$minor == "Endothelial-Kidney"] = "Stromal"
annot$major[annot$minor == "Fibroblast"] = "Stromal"
annot$major[annot$minor %in% c("Mast", "Monocyte")] = "Immune"

table(annot$major, annot$minor)
            
rownames(annot) = colnames(corMat)
annot$minor = as.factor(annot$minor)
annot$major = as.factor(annot$major)

my_col1 = colorRampPalette(RColorBrewer::brewer.pal(12, "Paired"))(length(unique(annot$major)))
names(my_col1) = levels(annot$major)
my_col2 = colorRampPalette(RColorBrewer::brewer.pal(12, "Paired"))(length(unique(annot$minor)))
names(my_col2) = levels(annot$minor)

my_col = list(major = my_col1, minor = my_col2)

hclust = hclust(as.dist(corMat), method = "single")
hclusters = cutree(hclust, k = 80)
hclusters = cutree(hclust, k = 400)
table(hclusters)

pdf(file = "test3.pdf", width = 25, height = 20)
idx_hclust = hclusters == names(which.max(table(hclusters)))
table(idx_hclust)
pheatmap::pheatmap(corMat[idx_hclust,idx_hclust], 
                   annotation_col = annot, 
                   annotation_colors = my_col,
                   show_rownames = FALSE, show_colnames = FALSE)
dev.off()
```

```{r}
cepoMat_combined = do.call(cbind, lapply(cepoMat2, function(x) x[idx,]))
colnames(cepoMat_combined) = paste0("_", rep(1:length(cepoMat2), sapply(cepoMat2, ncol)), "____", colnames(cepoMat_combined))
cepoMat_subset = cepoMat_combined[, !grepl("None|Neuronal", colnames(cepoMat_combined))]
cepoMat_subset = cepoMat_subset[, idx_hclust]

pList = lapply(c(1*10^-10), function(p_cut) {
  
  idx_gene = goi_mat$p < p_cut & goi_mat$z > 0
  goi1 = goi_mat$gene[idx_gene]
  idx_gene = goi_mat$p < p_cut & goi_mat$z < 0
  goi2 = goi_mat$gene[idx_gene]
  
  tmp = do.call(rbind, lapply(1:length(cepoMat), function(x) {

    tmp2 = grep(paste0("_", x, "____"), colnames(cepoMat_subset), value = TRUE)
    tmp2 = gsub(paste0("_", x, "____"), "", tmp2)
    idx_tmp = colnames(cepoMat[[x]]) %in% tmp2
    cepoMat_tmp = cepoMat[[x]][, idx_tmp, drop = FALSE]
    
    if (ncol(cepoMat_tmp) > 0) {
    
    mat = do.call(rbind, lapply(1:ncol(cepoMat_tmp), function(y) {
      
      
      geneStats = rank(cepoMat_tmp[,y])
      #geneIdx = intersect(names(goi_rank), names(test))
      
      p.in.up <- intersect(names(geneStats), goi1)
      p.out.up <- setdiff(names(geneStats), goi1)
      
      p.in.down <- intersect(names(geneStats), goi2)
      p.out.down <- setdiff(names(geneStats), goi2)
      
      wilcoxTest.mat <- matrix("NA", ncol = 4, nrow = 2)
      colnames(wilcoxTest.mat) <- c("cty", "pvalue", "# of substrates",
                                    "dir")
      
      wilcoxTest.mat[1, 1] <- colnames(cepoMat[[x]])[[y]]
      wilcoxTest.mat[1, 2] <- stats::wilcox.test(geneStats[p.in.up],
                                                 geneStats[p.out.up], alternative = "greater")$p.value
      wilcoxTest.mat[1, 3] <- length(p.in.up)
      wilcoxTest.mat[1, 4] <- "up"
      
      wilcoxTest.mat[2, 1] <- colnames(cepoMat[[x]])[[y]]
      wilcoxTest.mat[2, 2] <- stats::wilcox.test(geneStats[p.in.down],
                                                 geneStats[p.out.down], alternative = "greater")$p.value
      wilcoxTest.mat[2, 3] <- length(p.in.down)
      wilcoxTest.mat[2, 4] <- "down"
      
      
      res = as.data.frame(wilcoxTest.mat[,1:4, drop = FALSE])
      res$dataset = x
      res$condition = unique(sceList[[x]]$Condition)
      return(res)
      
    }))
    mat = as.data.frame(mat)
    mat$pvalue = as.numeric(mat$pvalue)
    return(mat)
          
    }
    
  }))
  
  matEnrich = tmp
  matEnrich$interaction = interaction(matEnrich$cty, matEnrich$condition)
  matEnrich$major = sapply(strsplit(matEnrich$cty, "___"), "[[", 1)
  matEnrich$minor = sapply(strsplit(matEnrich$cty, "___"), "[[", 2)
  
  #matEnrich$major[matEnrich$minor %in% c("T", "B", "NK", "Natural.Killer.T")] = "Lymphoid"
  matEnrich$logpvalue = -log10(matEnrich$pvalue)
  matEnrich = matEnrich[matEnrich$major %in% c("Immune", "Lymphoid", "Myeloid"), ] 
  matEnrich = matEnrich[!matEnrich$minor %in% c("Inner.Medullary.Collecting.Duct", "AT2", "Plasmacytoid.Dendritic", "Plasma"), ] 
  
  matEnrich$minor[matEnrich$minor %in% c("B")] = "B.cell.lineage"
  matEnrich$minor[matEnrich$minor %in% c("T")] = "T.cell.lineage"
  matEnrich$minor[matEnrich$minor %in% c("M2.Macrophage", "Macrophages")] = "Macrophage"
  matEnrich$minor[matEnrich$minor %in% c("Mast.cells")] = "Mast"
  matEnrich$minor[matEnrich$minor %in% c("Natural.Killer.T", 
                                         "Innate.lymphoid.cell.NK")] = "NK"
  matEnrich$minor[matEnrich$minor %in% c("Monocyte.cDC", "Classical.Dendritic", "Non.classical.monocyte")] = "Dendritic.cells"
  matEnrich$minor[matEnrich$minor %in% c("Monocyte.derived")] = "Monocytes"

  matEnrich$group = paste0(matEnrich$major, "_", matEnrich$minor, "_", matEnrich$dir)
    
  p = ggplot(matEnrich, aes(x = as.factor(dir), 
                            y = -log10(pvalue), fill = as.factor(condition), col = as.factor(condition))) + 
    #ggtitle(p_cut) + 
    geom_boxplot(show.legend = TRUE, notch = FALSE, alpha = 0.5) +
    #geom_jitter() + 
    ggsci::scale_fill_lancet() +
    ggsci::scale_color_lancet() +
    facet_wrap(~minor+major, ncol = 14) + 
    theme_bw()
  
  p1 = ggboxplot(matEnrich, x = "condition", y= "logpvalue", facet.by = "group", fill = "condition", ncol = 12,
            add = "jitter", palette = "npg") + 
    stat_compare_means(aes(method = "wilcox.test", label = paste0("p = ", after_stat(p.format))))
  
  return(p)
  
})

ptosave = pList[[1]]
ptosave %>% ggsave(filename = "boxplot_enrichment_minor_filtered.pdf", width = 15, height = 5)
```

### Figure 2G

```{r}
# You may need to change this working direction to the dataset you just downloaded.
setwd("~/Desktop/PhD/Pan Organ Allograft Dysfunction/Pan-Organ Allograft Dysfunction/Rejection/Normalised_Counts")
files <- list.files(pattern = "Blood")
Blood_Counts <- lapply(files, function(x){
  y <- read.csv(x) 
  y %>%
    dplyr::select(-X) %>%
    tibble::remove_rownames() %>%
    filter(!duplicated(Gene_Symbol) & !is.na(Gene_Symbol)) %>%
    tibble::column_to_rownames("Gene_Symbol") 
})

keep <- Blood_Counts %>% sapply(nrow) > 8000
Blood_Counts <- Blood_Counts[keep]
int <- Reduce(intersect, lapply(Blood_Counts, rownames))
Blood_Counts <- lapply(Blood_Counts, function(x)x[int,])
Counts_t <- lapply(Blood_Counts, t)

Blood_outcome <- lapply(Blood_Counts, function(x){
  factor(ifelse(grepl("Control", colnames(x)), "Control", "AR"))
})

Tissue <- sub(".csv", "", TOP:::str_split_n(files[keep], "_", 5))
Tissue[Tissue == ""] <- "Kidney"
```

```{r}
OrganSpecific_TOP <- function(OrganCounts, OrganOutcome){
  roc_perf <- list()
  for(i in seq_along(OrganCounts)){
    # Identify features changed in training sets. Reduces computational time. 
    sig.genes.Heart <- filterFeatures(x_list = OrganCounts[-i], y_list = OrganOutcome[-i], 
                                      contrast = "AR - Control", 
                                      combinationMethod = "Stouffer")
    
    # Subset by features from the training data.
    Data_temp = lapply(OrganCounts, "[", , sig.genes.Heart)
    
    # Convert to a matrix, while removing the testing set i
    x_list <- lapply(Data_temp, as.matrix)[-i]
    y_list <- lapply(OrganOutcome, factor)[-i]

    # Fit model on the training set.
    fCPOP_model <- TOP_model(x_list, y_list, sample_weights = TRUE, optimiseExponent = FALSE)
  
    # Predict on the testing set.
    pred <- predict_TOP(fCPOP_model$models, newx = Data_temp[[i]])

    # Calculate performance metrics for later.
    library(pROC)
    roc_perf[[i]] <- roc(factor(OrganOutcome[[i]]), pred)
  }
  return(unlist(lapply(roc_perf, function(x)as.numeric(x$auc))))
}
```


```{r}
PanOrgan_TOP <- function(Counts_t, Blood_outcome, Tissue, Tissue_idx){
  roc_perf <- list()
    for(i in Tissue_idx){
  
    # Identify features changed in training sets. Reduces computational time. 
    sig.genes.all <- filterFeatures(x_list = Counts_t, y_list = Blood_outcome, 
                                  contrast = "AR - Control", 
                                  combinationMethod = "Stouffer")
    
    # Subset by features from the training data.
    Data_temp = lapply(Counts_t, "[", , sig.genes.all)

    # Convert to a matrix, while removing the testing set i
    x_list <- lapply(Data_temp, as.matrix)[-i]
    y_list <- lapply(Blood_outcome, factor)[-i]

    tissue_weights = rep(Tissue[-i], lapply(y_list, length))

    # Fit model on the training set.
    fCPOP_model <- TOP_model(x_list, y_list, 
                           dataset_weights = tissue_weights,
                           sample_weights = TRUE, 
                           optimiseExponent = FALSE)
    pred <- predict_TOP(fCPOP_model$models, newx = Data_temp[[i]])

    # Calculate performance metrics for later.
    library(pROC)
    roc_perf[[i]] <- roc(factor(Blood_outcome[[i]]), pred)
    }
  roc_perf[!unlist(lapply(roc_perf, is.null))]
  return(unlist(lapply(roc_perf, function(x)as.numeric(x$auc))))
}
```

Let's predict the same datasets using all models.
```{r}
Heart <- Tissue == "Heart"
Heart_idx <- which(Tissue == "Heart")

Heart_Performance <- OrganSpecific_TOP(Counts_t[Heart], Blood_outcome[Heart])
PanHeart_Performance <- PanOrgan_TOP(Counts_t, Blood_outcome, Tissue, Heart_idx)
```

Let's merge the two
```{r}
Dataset <- TOP:::str_split_n(files[Heart], "_", 2)

names(Heart_Performance) <- Dataset
Heart_Performance <- Heart_Performance %>% 
  enframe() %>%
  mutate(Metric = "AUC", Model = "Heart")

names(PanHeart_Performance) <- Dataset
PanHeart_Performance <- PanHeart_Performance %>% 
  enframe() %>%
  mutate(Metric = "AUC", Model = "All")

Performance_Heart_models <- rbind(Heart_Performance, Full_Performance) %>%
  mutate(Organ = "Heart")
```

Go predict Heart datasets using only Kidney data. 
```{r}
Tissue <- sub(".csv", "", TOP:::str_split_n(files[keep], "_", 5))
Tissue[Tissue == ""] <- "Kidney"
Kidney <- Tissue == "Kidney"
Kidney_idx <- which(Tissue == "Kidney")

Kidney_Performance <- OrganSpecific_TOP(Counts_t[Kidney], Blood_outcome[Kidney])
PanKidney_Performance <- PanOrgan_TOP(Counts_t, Blood_outcome, Tissue, Kidney_idx)
```

```{r}
Dataset <- TOP:::str_split_n(files[which(Kidney)], "_", 2)

names(Kidney_Performance) <- Dataset
Kidney_Performance <- Kidney_Performance %>% 
  enframe() %>%
  mutate(Metric = "AUC", Model = "Kidney")

names(PanKidney_Performance) <- Dataset
PanKidney_Performance <- PanKidney_Performance %>% 
  enframe() %>%
  mutate(Metric = "AUC", Model = "All")

Performance_Kidney_models <- rbind(Kidney_Performance, PanKidney_Performance) %>%
  filter(name != "GSE120649") %>% # Avoid double dipping control patients  
  mutate(Organ = "Kidney")
```

Now we repeat for the liver datasets. 
```{r}
# You may need to change this working direction to the dataset you just downloaded.
setwd("~/Desktop/PhD/Pan Organ Allograft Dysfunction/Pan-Organ Allograft Dysfunction/Rejection/Normalised_Counts")
files <- list.files(pattern = "Blood")
Blood_Counts <- lapply(files, function(x){
  y <- read.csv(x) 
  y %>%
    dplyr::select(-X) %>%
    tibble::remove_rownames() %>%
    filter(!duplicated(Gene_Symbol) & !is.na(Gene_Symbol)) %>%
    tibble::column_to_rownames("Gene_Symbol") 
})
int <- Reduce(intersect, lapply(Blood_Counts, rownames))
Blood_Counts <- lapply(Blood_Counts, function(x)x[int,])
Counts_t <- lapply(Blood_Counts, t)

Blood_outcome <- lapply(Blood_Counts, function(x){
  factor(ifelse(grepl("Control", colnames(x)), "Control", "AR"))
})
```

Go predict Liver datasets using only Liver data. 
```{r}
Tissue <- sub(".csv", "", TOP:::str_split_n(files, "_", 5))
Tissue[Tissue == ""] <- "Kidney"
Liver <- Tissue == "Liver"

Liver_counts = Counts_t[Liver]
Liver_outcome = Blood_outcome[Liver]

roc_Liver <- list()
# Only two datasets, so a standard lasso is most comparable. 
for(i in seq_along(Liver_counts)){
  
  model <- glmnet::cv.glmnet(Liver_counts[[i]], Liver_outcome[[i]], family = "binomial", alpha = 1)
  pred <- predict(model, newx = Liver_counts[[-i]])

  library(pROC)
  roc_Liver[[i]] <- roc(factor(Liver_outcome[[-i]]), pred)
}
```

Let's predict the same datasets using all datasets to build the pan-model.
```{r}
Liver_idx <- which(Tissue == "Liver")
PanLiver_Performance <- PanOrgan_TOP(Counts_t, Blood_outcome, Tissue, Liver_idx)
```


```{r}
Dataset <- TOP:::str_split_n(files, "_", 2)
Dataset <- Dataset[Liver]
Dataset <- make_clean_names(Dataset)

Liver_Performance <- unlist(lapply(roc_Liver, function(x)as.numeric(x$auc)))
names(Liver_Performance) <- Dataset
Liver_Performance <- Liver_Performance %>% 
  enframe() %>%
  mutate(Metric = "AUC", Model = "Liver")

PanLiver_Performance <- unlist(lapply(roc_AllLiver, function(x)as.numeric(x$auc)))
names(PanLiver_Performance) <- Dataset
PanLiver_Performance <- PanLiver_Performance %>% 
  enframe() %>%
  mutate(Metric = "AUC", Model = "All")

Performance_Liver_models <- rbind(Liver_Performance, PanLiver_Performance) %>%
  mutate(Organ = "Liver") 
```

```{r}
perf_dfAll <- do.call("rbind", list(Performance_Heart_models, Performance_Kidney_models, Performance_Liver_models))
data_foo <- perf_dfAll %>% mutate(new_x = factor(interaction(Organ,Model), 
                                                 levels = c('Heart.Heart','Heart.All','Kidney.Kidney',
                                                            'Kidney.All', "Liver.Liver", "Liver.All")))

ggplot(data_foo, aes(x = new_x, y = value, fill = as.character(Model))) + 
  geom_boxplot() + theme_bw() + geom_line(aes(group = name), alpha = 0.3) + 
  scale_fill_manual(values = c("#ffa600", "#ff6361", "#0087c6", "#7a5195"), name = "Training\nDatasets") +
  geom_point(aes(group = name), shape = 21, color = 'black') + theme(axis.text.x = element_blank()) +
  xlab("") + ylab("AUC")

data_foo <- data_foo %>%
  group_by(new_x) %>%
  summarise(median = median(value)) %>%
  mutate(median = round(median, 2)) 
colnames(data_foo) <- c("Model", "Median AUC")
```

### Figure 2H

```{r}
setwd("~/Desktop/PhD/Pan Organ Allograft Dysfunction/Pan-Organ Allograft Dysfunction/Rejection/Normalised_Counts")
files <- list.files(pattern = "Blood")
Blood_Counts <- lapply(files, function(x){
  y <- read.csv(x) 
  y %>%
    dplyr::select(-X) %>%
    tibble::remove_rownames() %>%
    filter(!duplicated(Gene_Symbol) & !is.na(Gene_Symbol)) %>%
    tibble::column_to_rownames("Gene_Symbol") 
})

int <- Reduce(intersect, lapply(Blood_Counts, rownames))
Blood_Counts <- lapply(Blood_Counts, function(x)x[int,])
Counts_t <- lapply(Blood_Counts, t)

Blood_outcome <- lapply(Blood_Counts, function(x){
  factor(ifelse(grepl("Control", colnames(x)), "Control", "AR"))
})

Tissue <- sub(".csv", "", TOP:::str_split_n(files, "_", 5))
Tissue[Tissue == ""] <- "Kidney"
tissue_list <- rep(Tissue, lapply(Blood_outcome, length))
```


```{r}
library(TOP)

# Pull out the top 50 genes across datasets. 
sig.genes.all <- filterFeatures(x_list = Counts_t, y_list = Blood_outcome, 
                                  contrast = "AR - Control", 
                                  nFeatures = 50, combinationMethod = "Stouffer")
Data_temp = lapply(Counts_t, "[", , sig.genes.all)

x_list <- lapply(Data_temp, as.matrix)
y_list <- lapply(Blood_outcome, factor)
dataset_weights <- tissue_list # weights of the different organs. Makes a pan-organ model. 

# Fit model on the training set. Again, there was a merge issue so TOP_model has been rewritten above.
top_blood_model <- myTOP_model(x_list, y_list, dataset_weights = tissue_list, sample_weights = TRUE)
PanOrgan_coefPlot <- TOP_coefPlot(top_blood_model, nFeatures = 20)

# Next we will filter DGE$counts so that we can predict on the AUSCAD dataset.
filt_lcpm <- DGE$counts[sig.genes.all,]
rej_predictions <- TOP::predict_TOP(top_blood_model$models, newx = as.matrix(t(filt_lcpm)))

# Add the predictions back to the DGElist object.
DGE$samples$TOPrej_predictions <- as.numeric(rej_predictions)
```


# It will be interesting to see how our pan-organ model compares with one that was only built on kidney datasets. 
```{r}
# Lets just keep the datasets belonging to kidney transplant datasets.
kidney_idx <- Tissue == "Kidney"
filt_list <- Counts_t[kidney_idx]
y_filt <- Blood_outcome[kidney_idx]

# Filter to top 50 genes across datasets.
kidney_features <- TOP::filterFeatures(filt_list, y_filt, contrast = "AR - Control", nFeatures = 50)
kidney_counts = lapply(filt_list, "[", , kidney_features)

# Fit model on the training set. Again, there was a merge issue so TOP_model has been rewritten above.
kidney_model <- myTOP_model(kidney_counts, y_filt)
Kidney_coefPlot <- TOP_coefPlot(kidney_model, nFeatures = 20)

filt_lcpm <- DGE$counts[sig.genes.all,]
kidrej_predictions <- TOP::predict_TOP(kidney_model$models, as.matrix(t(filt_lcpm)))

# Add the predictions back to the DGElist object.
DGE$samples$KIDrej_predictions <- as.numeric(kidrej_predictions)
```

# We can also compare the clinical data to our above models. To do so we will fit an ordinary glm.
```{r}
# Let's fit a model to creatinine, egfr and albumin. 
clinical_model <- glm(factor(rej_3m) ~ cr_3m + egfr_3m + alb_3m, data=DGE$samples, family=binomial)
data = DGE$samples %>%
  dplyr::select(cr_3m, egfr_3m, alb_3m)

# Store our predictions.
clinical_predictions <- predict(clinical_model, data, type="response")

# Add the predictions back to the DGElist object.
DGE$samples$clinical_predictions <- as.numeric(clinical_predictions)
```

# I wanted to have some custom colours. So I rewrote the ROCplot function from TOP.
```{r}
myROC_Plot <- function (roc_list) 
{
    data.labels <- TOP:::extractAUC(roc_list)
    pROC::ggroc(roc_list, size = 1.5) + theme_bw() +
  scale_color_manual(name = "Model", values = c("#FF9DA7", "#0A87C6", "#FBA730"),
                     labels = data.labels$label_long) +
  geom_segment(aes(x = 1, xend = 0, y = 0, yend = 1), color = "grey50", linetype = "dashed") +
  theme(legend.title = element_text(size = 14)) +
  theme(legend.text = element_text(size = 12)) +
  ggtitle("")
}
```

# Plot the different ROC curves.
```{r}
n1 <- myROC_Plot(list(`Clinical Model` = roc(DGE$samples$rej_3m ~ DGE$samples$clinical_predictions)))

n2 <- myROC_Plot(list(`Clinical Model` = roc(DGE$samples$rej_3m ~ DGE$samples$clinical_predictions),
                  `Kidney Model` = roc(DGE$samples$rej_3m  ~ DGE$samples$KIDrej_predictions)))

n3 <- myROC_Plot(list(`Clinical Model` = roc(DGE$samples$rej_3m ~ DGE$samples$clinical_predictions),
                  `Kidney Model` = roc(DGE$samples$rej_3m  ~ DGE$samples$KIDrej_predictions),
                  `PanOrgan Model` = roc(DGE$samples$rej_3m  ~ DGE$samples$TOPrej_predictions)))
n3
```

